Computable Poincar\'e--Friedrichs constants for the $L^{p}$ de Rham complex over convex domains and domains with shellable triangulations

### 2024-10-04 main.tex
further polishment

\section{{Nonlinear and linear potentials, the Poincar\'e--Friedrichs and Poincar\'e inequalities, and {analytical} Poincar\'e--Friedrichs constants in the gradient case}} \label{section:poincare}

In this section, we wish to explain why we employ linear potential operators to estimate the Poincar\'e--Friedrichs constants and we also address the difference of the Poincar\'e--Friedrichs inequality~\eqref{math:PF_gradient_2} with the Poincar\'e inequality where the mean value of $u$ is fixed to zero. We then review Poincar\'e--Friedrichs inequalities in the gradient case setting of Section~\ref{section:intro_grad}, with the emphasis on know results on analytical upper bounds for the constants. In particular, we take the opportunity to discuss convex domains as a special case. This survey serves as a building block in constructing computable constants over triangulated domains in a combinatorial way below but we believe it is also of independent interest.

\subsection{Nonlinear and linear potentials}

In the Banach setting $p \neq 2$, finding a norm-minimizing potential such as $\Phi_{\grad}( \bff )$ of~\eqref{math:pot_grad_def} for a given gradient $\bff$ is a nonlinear operation. Its existence may actually not be guaranteed unless $1 < p < \infty$. 
However, any \textit{linear} potential operator $\overline\Phi_{\grad} : \nabla W^{1,p}(\Omega) \rightarrow L^{p}(\Omega)$
satisfying $\nabla \overline\Phi_{\grad}( \bff ) = \bff$ for any $\bff \in \nabla W^{1,p}(\Omega)$ has the operator norm that \cye{gives} the upper bound 
\begin{align}\label{math:PF_P}
    \mwl{C_{\grad,\Omega,p} = } 
    \max\limits_{ u \in W^{1,p}(\Omega) \setminus \bbR } 
    \frac{ \| \Phi_{\grad}(\nabla u) \|_{L^{p}(\Omega)} }{ \| \nabla u \|_{\fooL^{p}(\Omega)} }
    \leq 
    \max\limits_{ u \in W^{1,p}(\Omega) \setminus \bbR } 
    \frac{ \| \overline\Phi_{\grad}(\nabla u) \|_{L^{p}(\Omega)} }{ \| \nabla u \|_{\fooL^{p}(\Omega)} } \cye{=: C_{\varnothing,\Omega,p}}
    .
\end{align}
This explains our interest in the construction of {\em linear potential operators}.

\subsection{Poincar\'e--Friedrichs and Poincar\'e inequalities}

\cye{In the setting of Section~\ref{section:intro_grad}, when $\Omega$ is of finite measure,} a natural \cye{(and unique, since the gradient potential constraint $\nabla \overline \Phi_{\grad}( \bff ) = \bff$ determines $\overline \Phi_{\grad}( \bff )$ up to a constant)} choice \cye{for a linear potential operator} is the average-free potential operator $\overline\Phi_{\grad} : \nabla W^{1,p}(\Omega) \rightarrow L^{p}(\Omega)$ that is prescribed by
\begin{align*}
    \overline\Phi_{\grad}( \nabla u ) \cye{:}= u - u_{\Omega},
    \quad 
    \forall 
    u \in W^{1,p}(\Omega)
    ,
\end{align*}
where $u_\Omega$ denotes the average of $u$ over the domain $\Omega$, that is,
\begin{align*}
    u_{\Omega} := \vol(\Omega)^{-1} \int_{\Omega} u(x) \;dx.
\end{align*}
This potential operator can also implicitly be defined by
\begin{align*}
    \overline \Phi_{\grad}( \bff ) 
    := 
    \argmin\limits_{ \substack{ u \in W^{1,p}(\Omega) \\ \int_\Omega u = 0 } } \| \nabla u - \bff \|_{\fooL^{p}(\Omega)},
    \quad 
    \forall 
    \bff \in \nabla W^{1,p}(\Omega)
    .
\end{align*} 
Its operator norm is the best constant $C_{\varnothing,\Omega,p} > 0$ in the {\em Poincar\'e inequality} (or also Poincar\'e--Wirtinger inequality)
\begin{align}\label{math:P_gradient}
    \| u - u_{\Omega} \|_{L^{p}(\Omega)}
    \leq 
    C_{\varnothing,\Omega,p} \| \nabla u \|_{\fooL^{p}(\Omega)},
    \quad 
    \forall 
    u \in W^{1,p}(\Omega)
    . 
\end{align}
Thus, rephrasing~\eqref{math:PF_P} in terms of~\eqref{math:PF_gradient_2} and~\eqref{math:P_gradient}, 
\begin{align*}
    C_{\grad,\Omega,p} \leq C_{\varnothing,\Omega,p}.
\end{align*}

Towards a converse inequality, 
let us first observe that the average of any $u \in W^{1,p}(\Omega)$ with $p < \infty$ satisfies the bound 
\begin{align} \label{math:aver_norm}
    \| u_\Omega \|_{L^{p}(\Omega)}^{p}
    = 
    \int_{\Omega} \left( \vol(\Omega)^{-1} \int_{\Omega} |u(x)| \;dx \right)^{p}
    \leq 
    \int_{\Omega} \vol(\Omega)^{-1} \int_{\Omega} |u(x)|^{p} \;dx
    = 
    \| u \|_{L^{p}(\Omega)}^{p}
    .
\end{align}
Here, we have used H\"older's or Jensen's inequality. 
In the case $p = \infty$, any $u \in L^{\infty}(\Omega)$ satisfies $\| u_\Omega \|_{L^{\infty}(\Omega)} \leq \| u \|_{L^{\infty}(\Omega)}$. 
We conclude that taking the average is a projection within Lebesgue spaces with unit norm. 
The triangle inequality now shows that 
\begin{align*}
    \| u - u_\Omega \|_{L^{p}(\Omega)} 
    \leq
    2
    \| u \|_{L^{p}(\Omega)},
    \quad 
    \forall
    u \in L^{p}(\Omega)
    .
\end{align*}
\begin{align*}
    \color{Emerald}
    \| u - u_\Omega \|_{L^{p}(\Omega)}^{p}
    \leq 
    2^{p-1} 
    \| u \|_{L^{p}(\Omega)}^{p}
    +
    2^{p-1} 
    \| u_\Omega \|_{L^{p}(\Omega)}^{p}
    \leq 
    2^{p} 
    \| u \|_{L^{p}(\Omega)}^{p}
    .
\end{align*}
\mwl{Via an application of the triangle inequality, }Thus, the Poincar\'e--Friedrichs inequality implies the Poincar\'e inequality with 
\begin{align*}
    C_{\grad,\Omega,p} \leq 2 C_{\varnothing,\Omega,p}.
\end{align*}
In the special case $p=2$, taking the average is an orthogonal projection, 
and so this improves to $\| u - u_\Omega \|_{L^{2}(\Omega)} \leq \| u \|_{L^{2}(\Omega)}$ for any $u \in L^{2}(\Omega)$. 
Hence,
\begin{align*}
    C_{\varnothing,\Omega,2} = C_{\grad,\Omega,2}. 
\end{align*}
\mwl{
This improvement follows from the orthogonality of constant functions to average-free functions in $L^2(\Omega)$,
and it also follows from the projection estimate (see, e.g.,~\cite{xu2003some}).
Stern's generalized projection estimate~\cite[Theorem~4.1,Remark~5.1]{stern2015banach} implies improved estimate for all $L^{p}$ spaces with $1 \leq p \leq \infty$:
since taking the average is a the projection onto the constants functions with unit norm, it now follows that 
}
From~\eqref{math:aver_norm}, Stern's generalized projection estimate~\cite[Theorem~4.1, Remark~5.1]{stern2015banach} implies improved estimate for all $L^{p}$ spaces with $1 \leq p \leq \infty$ as 
\begin{align*}
    \| u - u_\Omega \|_{L^{p}(\Omega)}
    \leq 
    \min\left( 2, 2^{|2/p-1|} \right)
    \| u \|_{L^{p}(\Omega)}
    = 
    2^{|2/p-1|} 
    \| u \|_{L^{p}(\Omega)}
    ,
    \quad 
    \forall 
    u \in L^{p}(\Omega)
    .
\end{align*}
Here, we have used $1 \leq 2^{|2/p-1|} \leq 2$ for $1 \leq p \leq \infty$.
The upper bound is only attained in the limit cases $p = 1$ and $p = \infty$, and the lower bound is only attained if $p = 2$, where we reproduce the improved estimate for the Hilbert space case.
We thus conclude 
\begin{align}\label{math:p_vs_pf}
    C_{\grad,\Omega,p} \leq 2^{|2/p-1|} C_{\varnothing,\Omega,p}.
\end{align}
In summary, our notion of Poincar\'e--Friedrichs constant is equivalent to the notion of Poincar\'e constant in the literature, up to a numerical factor that only depends on $1 \leq p \leq \infty$ and that is at most $2$.

%%%%

\section{Old : Review of Poincar\'e--Friedrichs constants}\label{section:poincare}

\mwl{
In this section, we wish to explain why we employ linear potential operators to estimate the Poincar\'e--Friedrichs constants and we also address the difference of the Poincar\'e--Friedrichs inequality~\eqref{math:PF_gradient_2} with the Poincar\'e inequality where the mean value of $u$ is fixed to zero. We then review Poincar\'e--Friedrichs inequalities in the gradient case setting of Section~\ref{section:intro_grad}, with the emphasis on know results on analytical upper bounds for the constants. In particular, we take the opportunity to discuss bounded convex domains as a special case. 
This survey serves as a building block in constructing computable constants over triangulated domains in a combinatorial way below but we believe it is also of independent interest.

It is generally easier to obtain upper bounds Constructing and analyzing a potential operator is generally easier if the potential is linear. 
Constructing linear potential operators is generally easier 
We will extend that construction of the potential when we address Poincar\'e--Friedrichs inequalities for the curl and divergence operators.

% Our particular notion of Poincar\'e--Friedrichs inequality, which we interpret as an upper bound for the operator norm of a generalized inverse of the gradient is more natural for our discussion because later sections will develop analogous inequalities for the curl and the divergence operators. 
}

@@

Whenever $\calT$ is a simplicial complex, 
the set of $k$-dimensional simplices in $\calT$ is denoted as $\subsimplex_{{k}}(\calT)$. 
% Additionally, the notations $\Vertices(\calT) := \subsimplex_{0}(\calT)$ and $\Edges(\calT) := \subsimplex_{1}(\calT)$ refer to the vertices and edges of the simplicial complex, respectively. In practice, we not always distinguish between points and singleton simplices. 
% Similarly, if $\calT$ is an $n$-dimensional simplicial complex, we write $\Faces(\calT) := \subsimplex_{n-1}(\calT)$ for the faces (that is, codimension one members) of this triangulation.\footnote{
Similarly, the notations $\Vertices(\calT) := \subsimplex_{0}(\calT)$ and $\Faces(\calT) := \subsimplex_{n-1}(\calT)$ refer to the vertices and the faces (that is, codimension one members) of this triangulation.\footnote{
    Our use of the term \textit{face} as is common in the finite element literature~\cite{brenner2008mathematical}
    and is synonymous with \textit{facet} as used in the literature on polyhedral combinatorics~\cite{schrijver1998theory}.
    Notably, this terminology differs from the uses \textit{face} and \textit{facet} in the theory of polyhedra~\cite{ziegler1995lectures}.
}
% Given $T \in \calT$, we let $\supsimplex(\calT,T)$ be the set of simplices in $\calT$ that contain $T$,
% and we write $\supsimplex_{k}(\calT,T)$ for the set of $k$-dimensional simplices in $\supsimplex(\calT,T)$.\footnote{\mwl{MWL: It might be more economical to get rid of the supersimplex set, it is only used in this section to define patches.}}

@@

\begin{lemma}\label{lemma:poincarefriedrichsoverfacepatch}
    Let $\calT$ be a triangulation.
    Let $T_{1}, T_{2} \in \calT$ be two $n$-simplices whose intersection is a common facet $F := T_1 \cap T_2$. 
    Write $U := T_1 \cup T_2$. 
    If $p \in [1,\infty]$ and $u \in W^{1,p}(\Omega)$,
    then 
    \begin{align*}
        \min\limits_{ c \in \bbR }
        \| u - c \|_{L^{p}(U)}
        &
%         \leq 
%         2 \Ceins{n} C_{{\rm EFNT},p}
%         \max\left( 
%             \frac{ \vol(T_1) }{ \vol(T_2) }
%             ,
%             \frac{ \vol(T_2) }{ \vol(T_1) }
%         \right)^{\frac 1 p}
%         \\&\qquad\qquad 
%         \cdot 
%         \max\left( \diam(T_1), \diam(T_2) \right)
%         \| \nabla u \|_{L^{p}(U)}
%         \\&
        \leq 
        C_{{\PF},T_1 \cup T_2,p}
        \| \nabla u \|_{L^{p}(U)}
        .
    \end{align*}
    Here, $C_{{\PF},T_1 \cup T_2,p} = 2 \Ceins{n} C_{{\rm EFNT},p} \volumeratio(\calT)^{\frac 1 p} \max\left( \diam(T_1), \diam(T_2) \right)$. 
\end{lemma}
% \begin{remark}
%     The result is sharp when $T_1$ and $T_2$ have equal volume and diameter, and that $U$ is not necessarily convex. 
% \end{remark}
\begin{proof}
    Without loss of generality, $F$ has the vertices $v_0, \dots, v_{n-1}$,
    and $z_{1} \in T_{1}$ and $z_{2} \in T_{2}$ are the remaining vertices of the two triangles. 
    Let $\Delta_{1} = \Delta^{n}$ be the reference $n$-simplex and let $\Delta_{2}$ be obtained from it by flipping the $n$-th coordinate.
    We let $\varphi_{1} : \Delta_{1} \rightarrow T_{1}$ and $\varphi_{2} : \Delta_{2} \rightarrow T_{2}$
    be affine transformations 
    that map the origin to $v_0$, that map each unit vector $e_{i}$ to $v_{i}$ for $i = 1, \dots, n-1$,
    and that satisfy $\varphi_{1}(e_{n}) = z_{1}$ and $\varphi_{2}(-e_{n}) = z_{2}$.
    Write $\hat U := \Delta_1 \cup \Delta_2$.
    We have a bi-Lipschitz mapping $\varphi : \hat U \rightarrow U$.

@@

We now collect important examples of shellable triangulations. Essentially, in two space dimensions, interesting triangulations shellable, but starting from three space dimensions, non-shellable situations can arise. Our main interest are local patches (stars) within triangulations: these are shellable up to three space dimensions, but not necessarily beyond.

@@

We state the simple observation that local stars can be subdivided to vertex stars.
The shape regularity remains controlled in that process.

% \begin{lemma}\label{lemma:stardivision}
%     Let $\calT$ be an $n$-dimensional simplicial complex
%     such that $\calT = \patch_{\calT}(S)$ for some $S \in \calT$.
%     Let $z_S$ be the barycenter of $S$ 
%     and let $\calT^{\star}$ be the simplicial complex obtained from barycentric subdivision of $S$.
%     If $T^{\star} \in \calT^{\star}$ and $T \in \calT$ are both $n$-dimensional with $T^{\star} \subseteq T$,
%     then $\vol(T^{\star}) = \vol(T) / n$.
% \end{lemma}
\begin{lemma}\label{lemma:stardivision}
    Let $T$ be an $n$-dimensional simplex with a $k$-dimensional subsimplex $S$ and let $z_S$ be the barycenter of $S$. 
    Let $T'$ be one of the $n$-dimensional simplices obtained by splitting $T$ in accordance with the barycentric subdivision of $S$ at $z_{S}$.
    \begin{itemize}
        \item $\vol(T') = \vol(T) / (k+1)$. 
        \item The height vector of $v \in \Vertices(T) \setminus \Vertices(S)$ in $T'$ is the height vector $v$ in $T$.
        \item The height vector of $z_S$ in $T'$ is the height vector of the single vertex $v \in \Vertices(T) \setminus \Vertices(T')$ in $T$,
scaled by $(k+1)^{-1}$.
    \end{itemize}
\end{lemma}
\begin{proof}
    This is easy.
\end{proof}

\begin{figure}
    \begin{center}
    \begin{tikzpicture}
        % Coordinates of the triangle
        \coordinate (A) at (0,0); % Vertex A = v_0
        \coordinate (B) at (4,0); % Vertex B = v_1
        \coordinate (C) at (1,3); % Vertex C = v_2
        \coordinate (M) at ($(A)!0.5!(B)$); % Midpoint M of AB
        % Heights from each vertex
        \coordinate (H_A) at ($(C)!(A)!(B)$); % Height from A to BC
        \coordinate (H_B) at ($(C)!(B)!(A)$); % Height from B to AC
        \coordinate (H_C) at ($(A)!(C)!(B)$); % Height from C to AB
        \coordinate (H_M) at ($(B)!(M)!(C)$); % Height from M to BC
        % Drawing the triangle
        \draw[thick] (A) -- (B) -- (C) -- cycle;
        % Drawing the bisector from C to AB and the height vectors 
        %\draw[dashed] (C) -- (M);
        \draw[] (C) -- (M);
        \draw[dashed,->,line width=1.0pt,>=latex] (H_A) -- (A); % Height vector from A, original 
        \draw[dashed,->,line width=1.0pt,>=latex] (H_M) -- (M); % Height vector from A, original 
        \draw[dashed,->,line width=1.0pt,>=latex] (H_C) -- (C); % Height vector from C, unaffected
        % Marking the height points
        \node[above right] at (C) {$v_2$};  
        \node[below left] at (A) {$v_0$};
        \node[below right] at (B) {$v_1$};
        \node[below] at (M) {$z_S$};
    \end{tikzpicture}
    \end{center}
    \caption{Illustration of Lemma~\ref{lemma:stardivision}. The triangle $[ v_0, v_1, v_2 ]$ is bisected at the edge $[ v_0, v_1 ]$, leading to two new triangles. The height vector from $v_2$ in all three triangles remains the same. The height vector from $z_S$ in the new triangle $[z_S, v_1, v_2]$ is one half of the height vector from $v_0$ in the original triangle.}
\end{figure}

>>>>

\section{Review of Poincar\'e and Poincar\'e--Friedrichs inequalities}\label{section:poincare}

@@

Whenever $\calT$ is a simplicial complex, 
the set of $k$-dimensional simplices in $\calT$ is denoted as $\subsimplex_{{k}}(\calT)$. 
% Additionally, the notations $\Vertices(\calT) := \subsimplex_{0}(\calT)$ and $\Edges(\calT) := \subsimplex_{1}(\calT)$ refer to the vertices and edges of the simplicial complex, respectively. In practice, we not always distinguish between points and singleton simplices. 
% Similarly, if $\calT$ is an $n$-dimensional simplicial complex, we write $\Faces(\calT) := \subsimplex_{n-1}(\calT)$ for the faces (that is, codimension one members) of this triangulation.\footnote{
Similarly, the notations $\Vertices(\calT) := \subsimplex_{0}(\calT)$ and $\Faces(\calT) := \subsimplex_{n-1}(\calT)$ refer to the vertices and the faces (that is, codimension one members) of this triangulation.\footnote{
    Our use of the term \textit{face} as is common in classical geometry and the finite element literature~\cite{brenner2008mathematical}
    and is synonymous with \textit{facet} as used in the literature on polyhedral combinatorics~\cite{schrijver1998theory}.
    Notably, this terminology differs from the uses \textit{face} and \textit{facet} in the theory of polyhedra~\cite{ziegler1995lectures}.
}
% Given $T \in \calT$, we let $\supsimplex(\calT,T)$ be the set of simplices in $\calT$ that contain $T$,
% and we write $\supsimplex_{k}(\calT,T)$ for the set of $k$-dimensional simplices in $\supsimplex(\calT,T)$.\footnote{\mwl{MWL: It might be more economical to get rid of the supersimplex set, it is only used in this section to define patches.}}

@@

\begin{lemma}\label{lemma:poincarefriedrichsoverfacepatch}
    Let $\calT$ be a triangulation.
    Let $T_{1}, T_{2} \in \calT$ be two $n$-simplices whose intersection is a common face $F := T_1 \cap T_2$. 
    Write $U := T_1 \cup T_2$. 
    If $p \in [1,\infty]$ and $u \in W^{1,p}(\Omega)$,
    then 
    \begin{align*}
        \min\limits_{ c \in \bbR }
        \| u - c \|_{L^{p}(U)}
        &
%         \leq 
%         2 \Ceins{n} C_{{\rm EFNT},p}
%         \max\left( 
%             \frac{ \vol(T_1) }{ \vol(T_2) }
%             ,
%             \frac{ \vol(T_2) }{ \vol(T_1) }
%         \right)^{\frac 1 p}
%         \\&\qquad\qquad 
%         \cdot 
%         \max\left( \diam(T_1), \diam(T_2) \right)
%         \| \nabla u \|_{L^{p}(U)}
%         \\&
        \leq 
        C_{{\PF},T_1 \cup T_2,p}
        \| \nabla u \|_{L^{p}(U)}
        .
    \end{align*}
    Here, $C_{{\PF},T_1 \cup T_2,p} = 2 \Ceins{n} C_{{\rm EFNT},p} \volumeratio(\calT)^{\frac 1 p} \max\left( \diam(T_1), \diam(T_2) \right)$. 
\end{lemma}
% \begin{remark}
%     The result is sharp when $T_1$ and $T_2$ have equal volume and diameter, and that $U$ is not necessarily convex. 
% \end{remark}
\begin{proof}
    Without loss of generality, $F$ has the vertices $v_0, \dots, v_{n-1}$,
    and $z_{1} \in T_{1}$ and $z_{2} \in T_{2}$ are the remaining vertices of the two triangles. 
    Let $\Delta_{1} = \Delta^{n}$ be the reference $n$-simplex and let $\Delta_{2}$ be obtained from it by flipping the $n$-th coordinate.
    We let $\varphi_{1} : \Delta_{1} \rightarrow T_{1}$ and $\varphi_{2} : \Delta_{2} \rightarrow T_{2}$
    be affine transformations 
    that map the origin to $v_0$, that map each unit vector $e_{i}$ to $v_{i}$ for $i = 1, \dots, n-1$,
    and that satisfy $\varphi_{1}(e_{n}) = z_{1}$ and $\varphi_{2}(-e_{n}) = z_{2}$.
    Write $\hat U := \Delta_1 \cup \Delta_2$.
    We have a bi-Lipschitz mapping $\varphi : \hat U \rightarrow U$.

@@

We now collect important examples of shellable triangulations. Essentially, in two space dimensions, interesting triangulations are shellable, but starting from three space dimensions, non-shellable situations can arise. Our main interest are local patches (stars) within triangulations: these are shellable up to three space dimensions, but not necessarily beyond.

@@

We will use the following observaton, which we state without proof, 
that controls the volume and some heights when a simplex is partitioned via barycentric subdivision.

% \begin{lemma}\label{lemma:stardivision}
%     Let $\calT$ be an $n$-dimensional simplicial complex
%     such that $\calT = \patch_{\calT}(S)$ for some $S \in \calT$.
%     Let $z_S$ be the barycenter of $S$ 
%     and let $\calT^{\star}$ be the simplicial complex obtained from barycentric subdivision of $S$.
%     If $T^{\star} \in \calT^{\star}$ and $T \in \calT$ are both $n$-dimensional with $T^{\star} \subseteq T$,
%     then $\vol(T^{\star}) = \vol(T) / n$.
% \end{lemma}
\begin{lemma}\label{lemma:stardivision}
    Let $T$ be an $n$-dimensional simplex with a $k$-dimensional subsimplex $S$ and let $z_S$ be the barycenter of $S$. 
    Let $T'$ be one of the $n$-dimensional simplices obtained by splitting $T$ in accordance with the barycentric subdivision of $S$ at $z_{S}$.
    \begin{itemize}
        \item $\vol(T') = \vol(T) / (k+1)$. 
        \item The height vector of $v \in \Vertices(T) \setminus \Vertices(S)$ in $T'$ is the height vector $v$ in $T$.
        \item The height vector of $z_S$ in $T'$ is the height vector of the single vertex $v \in \Vertices(T) \setminus \Vertices(T')$ in $T$,
scaled by $(k+1)^{-1}$.
	\hfill\qed
    \end{itemize}
\end{lemma}

\begin{figure}
    \begin{center}
    \begin{tikzpicture}
        % Coordinates of the triangle
        \coordinate (A) at (0,0); % Vertex A = v_0
        \coordinate (B) at (4,0); % Vertex B = v_1
        \coordinate (C) at (1,3); % Vertex C = v_2
        \coordinate (M) at ($(A)!0.5!(B)$); % Midpoint M of AB
        % Heights from each vertex
        \coordinate (H_A) at ($(C)!(A)!(B)$); % Height from A to BC
        \coordinate (H_B) at ($(C)!(B)!(A)$); % Height from B to AC
        \coordinate (H_C) at ($(A)!(C)!(B)$); % Height from C to AB
        \coordinate (H_M) at ($(B)!(M)!(C)$); % Height from M to BC
        % Drawing the triangle
        \draw[thick] (A) -- (B) -- (C) -- cycle;
        % Drawing the bisector from C to AB and the height vectors 
        %\draw[dashed] (C) -- (M);
        \draw[] (C) -- (M);
        \draw[dashed,->,line width=1.0pt,>=latex] (H_A) -- (A); % Height vector from A, original 
        \draw[dashed,->,line width=1.0pt,>=latex] (H_M) -- (M); % Height vector from A, original 
        \draw[dashed,->,line width=1.0pt,>=latex] (H_C) -- (C); % Height vector from C, unaffected
        % Marking the height points
        \node[above right] at (C) {$v_2$};  
        \node[below left] at (A) {$v_0$};
        \node[below right] at (B) {$v_1$};
        \node[below] at (M) {$z_S$};
    \end{tikzpicture}
    \begin{tikzpicture}

        % Coordinates of the triangle
        \coordinate (A) at (0,0); % Vertex A = v_0
        \coordinate (B) at (4,0); % Vertex B = v_1
        \coordinate (C) at (1,3); % Vertex C = v_2
        % Barycenter (centroid) of the triangle
        \coordinate (G) at ($(A)!1/3!(B)!1/3!(C)$);
        % Drawing the triangle
        \draw[thick] (A) -- (B) -- (C) -- cycle;
        % Tripartitioning the triangle: draw lines from vertices to the barycenter
        \draw[] (A) -- (G);
        \draw[] (B) -- (G);
        \draw[] (C) -- (G);
        % Heights from the barycenter to the sides of the triangle
        \coordinate (H_GA) at ($(B)!(G)!(C)$); % Height from G to BC
        \coordinate (H_GB) at ($(A)!(G)!(C)$); % Height from G to AC
        \coordinate (H_GC) at ($(A)!(G)!(B)$); % Height from G to AB

        % Drawing the height vectors from the barycenter
        \draw[dashed,->,line width=1.0pt,>=latex] (H_GA) -- (G); % Height from G to BC
        \draw[dashed,->,line width=1.0pt,>=latex] (H_GB) -- (G); % Height from G to AC
        \draw[dashed,->,line width=1.0pt,>=latex] (H_GC) -- (G); % Height from G to AB

        % Marking the points
        \node[above right] at (C) {$v_2$};  
        \node[below left] at (A) {$v_0$};
        \node[below right] at (B) {$v_1$};
        \node[right] at (G) {$z_T$};
    \end{tikzpicture}
    \end{center}
    \caption{Illustration of Lemma~\ref{lemma:stardivision}. Left: the triangle $T = [ v_0, v_1, v_2 ]$ is bisected at the edge $S = [ v_0, v_1 ]$, leading to two new triangles. The height vector from $v_2$ in all three triangles remains the same. The height vector from $z_S$ in the new triangle $[z_S, v_1, v_2]$ is one half of the height vector from $v_0$ in the original triangle. Right: the triangle $T$ is trisected, leading to three new triangles. The height of $z_T$ to from some edge is one third of the original height of that edge.}
\end{figure}
